---
title: "simpleRegress"
author: "Louie A"
date: "7/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
The nick of th bear is at 20 and the predicted length is around 57. Looking at the data, the residual standard error is at 5.376 and the R-squared is at 0.7476.

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 




```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```
The residual standard error for this one is at 4.939. It is a better model than the previous 5.376 and R-squared for thi one is 0.7869 in which it is higher the 0.7476.




```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

In the third graph the residual standard erorr is at 7.32 and t r-squared is 0.5409. in the comparision this result ranks last on the list





















