---
title: "simpleRegress"
author: "Louie A"
date: "7/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
The nick of th bear is at 20 and the predicted length is around 57. Looking at the data, the residual standard error is at 5.376 and the R-squared is at 0.7476.

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 




```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```
The residual standard error for this one is at 4.939. It is a better model than the previous 5.376 and R-squared for thi one is 0.7869 in which it is higher the 0.7476.




```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

In the third graph the residual standard erorr is at 7.32 and t r-squared is 0.5409. in the comparision this result ranks last on the list


```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)

```
In WEIGHT the residual standard error is 5.432 and R-squared is 0.7472. This would be as of now would be second best. the order of residual standard error being CHEST 4.939, WEIGHT5.432 NECK 5.736, and HEADWTH 7.32. While R-squared by now is HEADWYH 0.5409, WEIGHT .07432, NECK 0.7467 and CHEST 0.791. 





```{r}
MYHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(MYHEADLEN)
summary.lm(MYHEADLEN)

```
The residual standard error is 4.235 and the r-squared 0.8463. as of now this has the worst r- squared and residual standard error .The new list for residual being  4.235 HEADLEN, CHEST 4.939, WEIGHT5.432 NECK 5.736, and HEADWTH 7.32. While R-squared by now is HEADWTH 0.5409, WEIGHT .07432, NECK 0.7467 , CHEST 0.791 and HEADLEN 0.8434. 




```{r}
MYSEX <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
abline(MYSEX)
summary.lm(MYSEX)

```
Sex is going to have a different type of graph but still has the same results we can use. SEX has both the best and worst numbers in each category.   The new list for residual being  4.235 HEADLEN, CHEST 4.939, WEIGHT5.432 NECK 5.736, HEADWTH 7.32 and Sex 10.75. While R-squared by now is SEX 0.01065, HEADWTH 0.5409, WEIGHT .07432, NECK 0.7467 , CHEST 0.791 and HEADLEN 0.8434. 



```{r}
MYMONTH <- lm(LENGTH~MONTH,data=bears)
plot(LENGTH~MONTH,data=bears)
abline(MYMONTH)
summary.lm(MYMONTH)
```

Month now takes the top with residual standard error being 10.99 and r-squared being 0.08423. The new list for residual being  4.235 HEADLEN, CHEST 4.939, WEIGHT 5.432 NECK 5.736, HEADWTH 7.32, Sex 10.75 and MONTH 10.99. While R-squared by now is SEX 0.01065, HEADWTH 0.5409, WEIGHT .07432 MONTH 0.08423, NECK 0.7467 , CHEST 0.791 and HEADLEN 0.8434. 



Conclusion.
With all the data grafted out and organized the the order for residual standard square is HEADLEN 4.235 , CHEST 4.939, WEIGHT 5.432 NECK 5.736, HEADWTH 7.32, Sex 10.75 and MONTH 10.99. while list for r-squared is 
SEX 0.01065, WEIGHT 0.07432 ,MONTH 0.08423, HEADWTH 0.5409 , NECK 0.7467 , CHEST 0.791 and HEADLEN 0.8434.



















